\addchap{Acknowledgements}

I would like to thank everybody in Tanaka-lab. The working atmosphere here in perfect, generous and warm. 
\begin{itemize}
	\item My advisor Pr.~Hajime Tanaka gave me support and constant attention during my years in his lab. I thank him especially for his open mind, humanity and unmovable smile.
	\item Special thanks to Mika Kobayashi who helped me for thousands of technical problems and endured being my neighbour for three years with kindness and friendship. 
	\item Many thanks to Mathieu Nespoulous for the long conversations in French. I already miss them. 
	\item Of course, thanks to Paddy Royall, emeritus member of the lab from who I learned everything I now about confocal microscopy and who handed me my first particle tracking software. Without you, my bibliography would be half what it is.
	\item Rebecca Rice did the \ac{SEM} imaging and sizing at the University of Bristol.
	\item I learned from Didi Derks the method to synthesize fluorescently labeled and sterically stabilized \ac{PMMA} colloids.
	\item The colloids I finally used were synthesised by Hideo Tsurusawa.
	\item Alex Malins proofread my English. I thank him for all the useful discussions about structure identification.
	\item I would not have chosen this theme for my thesis without the merriness of Takeshi Kawasaki and his communicative enthusiasm. I owe him so much in discussions, insights, 3 AM mail answers and friendship that I will run out of paper before ending the list. I thank him especially for lending me his precious data.
	\item Many thanks to Satoko Oizumi and Kumiko Kamata for their kind help in anything related to organisation.
\end{itemize}

I send my best wishes to all the people very far from me, that I met only by the Internet for some of them, and who helped me and supported me online.
\begin{itemize}
	\item Christophe Philemotte, Jean-Michel Deruty and Frank Gosselin literally taught me how to program, more than any computer science teacher I ever had.
	\begin{itemize}
		\item Toch, I am eager to meet you soon. Thanks for showing me the way through the thesis.
		\item Sogarf, thanks to read all my posts on the forum. Thanks for everyday's "Salut Iess" that cheered me up.
	\end{itemize}
	\item Olivier Le Floch told me by the way "You should look into spatial indexing" and saved me hundred of hours of machine time.
	\item Dustin Spicuzza kindly answered my questions on his blog and helped me to fit his RTree library to my needs.
	\item All the peoples from Ideo and Cur-Asso who supported my dreams.
\end{itemize}

Benoît, Pierrick, Isabelle and Gaël, I really miss your regular presence and conversations. I am so glad of your inalterable friendship each time we meet, even after years of separation. What are you waiting for to visit Japan?

Last but not least, I want to thank my families in France and in Japan.
\begin{itemize}
	\item My family in France that I miss so much.
	\item My family in Japan who offered me the perfect environment and mood to write. Special thanks to the cooking of Masako. Masanori, take care of you.
	\item My wife Lisa and my daughter Meï who who enlighten each day of my life and supported me all the way through. With love.
\end{itemize}

\addchap{Colophon}

This thesis is typed in \XeTeX, a flavour of \TeX.

Fonts are \emph{TeX Gyre Pagella} for body text and headings, \emph{Asana-Math} for mathematical expressions and \emph{IPA Mincho} for Japanese headings.

The figures are produced by \emph{Gnuplot} (2D), \emph{Paraview} (3D) with the help of \emph{Inkscape} and \emph{the GIMP}.

The \emph{C++} and \emph{Python} code used in this thesis is available online~\citep{LeocmachColloids}.